---
title: "Baseline biomarker barplots (TMB, PD-L1, MSI/dMMR"
output: html_document
date: "2024-05-23"
---

# Set up

### Libraries

```{r}
library(here)
library(tidyverse)
library(grid)
library(ggbreak) 
library(ggplot2)
library(ggpubr)
library(svglite)
library(rstatix)
library(Cairo)
library(openxlsx)
library(readxl)
```

### Directory pathing

```{r}
here::i_am("code/02.4_aging_baseline.tmb.pdl1.msi.dmmr_analysis.Rmd")

output.path.rds <- here("output/RDS files")
if (!dir.exists(here(output.path.rds))) dir.create(here(output.path.rds))

output.path.excel <- here("output/Excel files")
if (!dir.exists(here(output.path.excel))) dir.create(here(output.path.excel))

output.path.figure_files <- here("output/Figure files")
if (!dir.exists(here(output.path.figure_files))) dir.create(here(output.path.figure_files))

output.path.tiff <- here("output/Figure files/Tiff files")
if (!dir.exists(here(output.path.tiff))) dir.create(here(output.path.tiff))

output.path.svg <- here("output/Figure files/SVG files")
if (!dir.exists(here(output.path.svg))) dir.create(here(output.path.svg))

output.path.pdf <- here("output/Figure files/PDF files")
if (!dir.exists(here(output.path.pdf))) dir.create(here(output.path.pdf))

```

### Source functions

```{r}
source(here("code/functions/function_boxplot.R"))

```

### Import data

```{r}

clinical <- readRDS(here("output/RDS files/clinical.rds"))

file_path <- here("data/clinical/Aging_deid_TMB.PDL1.MSIH.dMMR_3.7.24_data.extracted.xlsx")

biomarker_data <- read_excel(file_path) %>%
  select(!pub_id:`pub_id paper`) %>%
  rename(TMB = "TMB (mut/Mb)",
         PDL1_score_type = "PDL1 score type (i.e. TPS, CPS, etc)",
         `MSI-H.dMMR_status` = "MSI-H/dMMR (yes or no)")
```

### Combine data

```{r}

clinical_biomarker_tb <- clinical %>%
  left_join(biomarker_data, by = "pt id") %>%
  select(`pt id`, `pub_id paper`, `age on study`, age65_status, TMB:`MSI-H.dMMR_status`)

```

### Data processing

```{r}
TMB_data <- clinical_biomarker_tb %>%
  filter(!is.na(TMB))

PDL1_data <- clinical_biomarker_tb %>%
  filter(!is.na(`PDL1 expression classification`))

MSI.dMMR_data <- clinical_biomarker_tb %>%
  filter(!is.na(`MSI-H.dMMR_status`))

MSI.dMMR_data$`MSI-H.dMMR_status` <- factor(MSI.dMMR_data$`MSI-H.dMMR_status`, levels = c("Yes","No"))
```


colors <- c(pal_npg()(10))
color_from_viridis <- viridis(1, option = "D", alpha = 1)
color_pal_v1 <- c("#284B96",
                 "#4AB8A7",
                 "#F49835",
                 "#878787",
                 "#544667",
                 "#C37672")

color_pal_v2 <- c(colors[2], color_from_viridis)

### Color palettes

```{r}

color_pal_nejm <- c("#0072B5FF","#BC3C29FF")

three_colors_bar <- c("#d1495b","#edae49","#00798c")

two_colors_bar <- c("#d1495b","#00798c")
```

### Special characters for graphs

```{r}
# For >= symbol in strings
greater_than_equal <- "\u2265"
less_than <- "\u003c"

age_greater.equal_65 <- paste0("Age ",greater_than_equal,"65")

age_less_65 <- paste0("Age ",less_than,"65")

age_greater.equal_75 <- paste0("Age ",greater_than_equal,"75")

age_less_75 <- paste0("Age ",less_than,"75")

```

# Statistics

# TMB t.test table

```{r}
# Wilcoxon table cytokines age 65 status (cytokines)

TMB_data %>%
  group_by(age65_status) %>%
  summarise(count = n())

tmb_age65status_t.test_table <- TMB_data  %>%
  summarise(category = "TMB",
            mean_age65.older = mean(TMB[age65_status == "yes"]),
            mean_age65.younger = mean(TMB[age65_status == "no"]),
            std_age65.older = sd(TMB[age65_status == "yes"]),
            std_age65.younger = sd(TMB[age65_status == "no"]),
            pval = t.test(TMB ~ age65_status, paired = FALSE)$p.value) 

```

### PDL1 count information

```{r}

PDL1_data %>%
  group_by(age65_status) %>%
  summarise(count = n())

```

### PDL1 Fisher

```{r}
# Fisher
fisher.PDL1.age.tb <- PDL1_data %>%
  select(age65_status, `PDL1 expression classification`) %>%
  group_by(age65_status, `PDL1 expression classification`) %>%
  summarise(count = n())

fisher.PDL1.age.manual.df <- data.frame("no" = fisher.PDL1.age.tb$count[1:3], 
                                                 "yes" = fisher.PDL1.age.tb$count[4:6], 
                                                 row.names = c("high","low","none")
                                        )

fisher.test(fisher.PDL1.age.manual.df)
```

### MSI/dMMR count information

```{r}

MSI.dMMR_data %>%
  group_by(age65_status) %>%
  summarise(count = n())

```

### MSI/dMMR Fisher

```{r}
# Fisher
fisher.MSI.dMMR.age.tb <- MSI.dMMR_data %>%
  select(age65_status, `MSI-H.dMMR_status`) %>%
  group_by(age65_status, `MSI-H.dMMR_status`) %>%
  summarise(count = n())

fisher.MSI.dMMR.age.manual.df <- data.frame("no" = fisher.MSI.dMMR.age.tb$count[1:2], 
                                                 "yes" = fisher.MSI.dMMR.age.tb$count[3:4], 
                                                 row.names = c("high","none")
                                        )

fisher.test(fisher.MSI.dMMR.age.manual.df)
```

# Plots

### TMB boxplot

```{r}

TMB_plot <- TMB_data %>%
  ggplot(aes(x= age65_status, y= as.numeric(TMB), color = age65_status)) + 
    scale_color_manual(values = color_pal_nejm) + 
      geom_boxplot(size = 0.75, 
                 fill = 'white'
    ) +
    geom_jitter(alpha = 0.75, pch=19, size=1.5, position=position_jitterdodge(dodge.width=1, jitter.width=0.5)) +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0,22, by = 5)) +
  #scale_x_discrete(labels = c(age_less_65, age_greater.equal_65)) +
  scale_x_discrete(labels = c("Young", "Aged")) +
  labs(title = "TMB", 
       y = "TMB (mut/Mb)") +
  #stat_compare_means(method = "t.test") +
  stat_compare_means(comparisons = list(c("no","yes")), aes(label = after_stat(p.signif)), bracket.size = 0.5, label.y = 21, size = 5, method = "t.test") +
  theme(axis.text.y = element_text(color="black", size=12, angle = 90, hjust = 0.5), 
        axis.text.x = element_text(color="black",size=12, angle = 0), 
        axis.line = element_line(colour = "black"), 
        axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 14),
        plot.title = element_text(hjust = 0.5, size = 14),
        strip.text = element_blank(), strip.background = element_blank(),
        axis.text.y.left = element_text(angle = 0), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position = 'none')
  
print(TMB_plot)

```

### PDL1 Stacked Barplot

```{r}

PDL1.age_graph <- PDL1_data %>%
  count(`PDL1 expression classification`, age65_status) %>%
  group_by(age65_status) %>%
  mutate(pct= prop.table(n)*100) %>%
  ggplot() +
  aes(age65_status, pct, fill = `PDL1 expression classification`) +
  scale_fill_manual(name = "PD-L1 classification", values=three_colors_bar, labels = c("High","Low","None")) +
  geom_bar(stat="identity") +
  ylab("Patients (%)") +
  geom_text(aes(label=paste0(n, "(",sprintf("%1.1f", pct),"%)")),size = 6, color = "white",
            position=position_stack(vjust=0.5)) +
  theme_classic() +
  labs(#title = "Response by age status", 
       x = #"Age status",
       "") +
  scale_x_discrete(labels = c("Young", "Aged")) +
  #scale_x_discrete(labels = c(age_less_65, age_greater.equal_65)) +
  theme(plot.title = element_text(hjust = 0.75, size = 20), 
        axis.text.y.left = element_text(angle = 0), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.text.y = element_text(color="black", size=20, angle = 90, hjust = 0.5),
        axis.text.x = element_text(color="black",size=20, angle = 0), 
        axis.line = element_line(colour = "black"), 
        axis.title.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18) 
        ) 

print(PDL1.age_graph)

```


### MSI-H dMMR Stacked Barplot

```{r}

MSI.dMMR.age_graph <- MSI.dMMR_data %>%
  count(`MSI-H.dMMR_status`, age65_status) %>%
  group_by(age65_status) %>%
  mutate(pct= prop.table(n)*100) %>%
  ggplot() +
  aes(age65_status, pct, fill = `MSI-H.dMMR_status`) +
  scale_fill_manual(name = "MSI-H/dMMR\nstatus", values=two_colors_bar, labels = c("Yes","No")) +
  geom_bar(stat="identity") +
  ylab("Patients (%)") +
  geom_text(aes(label=paste0(n, "(",sprintf("%1.1f", pct),"%)")),size = 6, color = "white",
            position=position_stack(vjust=0.5)) +
  theme_classic() +
  labs(#title = "Response by age status", 
       x = #"Age status",
       "") +
  #scale_x_discrete(labels = c(age_less_65, age_greater.equal_65
  scale_x_discrete(labels = c("Young", "Aged")) +
  theme(plot.title = element_text(hjust = 0.75, size = 20), 
        axis.text.y.left = element_text(angle = 0), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(),
        axis.text.y = element_text(color="black", size=20, angle = 90, hjust = 0.5),
        axis.text.x = element_text(color="black",size=20, angle = 0), 
        axis.line = element_line(colour = "black"), 
        axis.title.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 18) 
        ) 

print(MSI.dMMR.age_graph)

```

### Plot saving

```{r}
ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Stacked_Barplot_MSI.dMMR_age.65.status.pdf"), 
       plot = MSI.dMMR.age_graph, 
       height = 7, width = 7,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Stacked_Barplot_PDL1_age.65.status.pdf"), 
       plot = PDL1.age_graph, 
       height = 7, width = 7,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Boxplot_TMB_age.65.status.pdf"), 
       plot = TMB_plot, 
       height = 3, width = 3,
       device = cairo_pdf,
       dpi = 600)

```

### Generation of supporting data file

```{r}
biomarker_supporting.data <- clinical_biomarker_tb %>%
  select(`pub_id paper`,age65_status,TMB,`PDL1 expression classification`,`MSI-H.dMMR_status`)

write.xlsx(biomarker_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "Aging_TMB.PDL1.MSI.dMMR_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)

```

