---
title: "Cytokine heatmaps"
output: html_document
date: "2024-05-23"
---
# Set up

### Libraries

```{r, results = 'hide', warning = FALSE, message = FALSE}
# Libraries
library(here)
library(tidyverse)
library(ggplot2)
library(pheatmap)
library(viridis)
library(ggpubr)
library(gridExtra)
library(wesanderson)
library(grid)
library(ggbreak) 
library(ggsci)
library(ComplexHeatmap)
library(khroma)
library(greekLetters)
library(Cairo)
library(RColorBrewer)
library(openxlsx)
library(readxl)
library(circlize)

```

### Directory pathing

```{r, results = 'hide', warning = FALSE, message = FALSE}
here::i_am("code/02.6_aging_cyt_heatmaps.Rmd")

output.path.rds <- here("output/RDS files")
if (!dir.exists(here(output.path.rds))) dir.create(here(output.path.rds))

output.path.excel <- here("output/Excel files")
if (!dir.exists(here(output.path.excel))) dir.create(here(output.path.excel))

output.path.figure_files <- here("output/Figure files")
if (!dir.exists(here(output.path.figure_files))) dir.create(here(output.path.figure_files))

output.path.tiff <- here("output/Figure files/Tiff files")
if (!dir.exists(here(output.path.tiff))) dir.create(here(output.path.tiff))

output.path.svg <- here("output/Figure files/SVG files")
if (!dir.exists(here(output.path.svg))) dir.create(here(output.path.svg))

output.path.pdf <- here("output/Figure files/PDF files")
if (!dir.exists(here(output.path.pdf))) dir.create(here(output.path.pdf))
```

### Import files

```{r}
clinical_tb <- readRDS(here("output/RDS files/clinical.rds"))

clinical_cytokine_baseline <- readRDS(here("output/RDS files/demo_baseline_log2.rds"))

clinical_cytokine_on.treatment_change <- readRDS(here("output/RDS files/demo_change.rds"))

# baseline tibble
clinical_cytokine_baseline_age.status <- clinical_cytokine_baseline %>%
  ungroup() %>%
  distinct(`pt id`, cytokines, .keep_all = TRUE)

# cytokine timepoint 1 tibble
clinical_change_timepoint1_age.status <- clinical_cytokine_on.treatment_change %>%
  arrange(`pt id`,`blood timepoint order`) %>%
  distinct(`pt id`, cytokines, .keep_all = TRUE) 

```

### Special characters for graphs

```{r}
# For >= symbol in strings
greater_than_equal <- "\u2265"
less_than <- "\u003c"

age_greater.equal_65 <- paste0("Age ",greater_than_equal,"65")

age_less_65 <- paste0("Age ",less_than,"65")
```

### Color palettes

```{r}
colors_npg <- pal_npg()(10)

colors_anno_label_heatmap <- c("#544667","#C37672","#F49835","#4AB8A7","#284B96","#878787")

colors_cyt_label_heatmap <- c("#544667","#C37672","#F49835","#4AB8A7","#284B96","#878787")

color_pal_blu.red.yellow <- c("#084c61","#db504a","#e3b505","#4f6d7a","#56a3a6")

colors <- c(pal_npg()(10))
color_from_viridis <- viridis(1, option = "D", alpha = 1)
color_pal_w.viridis <- c(colors[2], color_from_viridis)

color_pal <- c("#15607a","#3cc4ad","#db504a","#F49835","#594F4F")
```

# Data set up and processing

## Cytokine order and groupings in heatmap

```{r}
options(scipen=999)

#Cytokine order
cyt_order <- c('il-2', 'il-12(p70)', 'il-12(p40)', 'ifng', 'tnfa',
               'il-4', 'il-5', 'il-9', 'il-13', 'il-25',
               'il-6', 'il-17a', 'il-17f', 'il-21', 'il-22', 'il-23',
               'il-10', 'il-35',
               'il-8', 'itac', 'mcp-1', 'mip-1a', 'mip-1b', 'rantes', 'mip-3a', 'mpif-1', 'mig','ip-10',
               'il-1a', 'il-1b', 'il-1ra', 'il-3','il-15', 'il-18', 'g-csf', 'gm-csf', 'baff', 'vegf-a','scd40l'
)

#Cytokine groupings
chemokines <- c('il-8', 'itac', 'mcp-1', 'mip-1a', 'mip-1b', 'rantes', 'mip-3a', 'mpif-1', 'mig', 'ip-10')
other <- c('il-1a', 'il-1b', 'il-1ra', 'il-3', 'il-15', 'il-18', 'g-csf', 'gm-csf', 'baff', 'vegf-a', 'scd40l')
th2 <- c('il-4', 'il-5', 'il-9', 'il-13', 'il-25')
treg <- c('il-10', 'il-35')
th17 <- c('il-6', 'il-17a', 'il-17f', 'il-21', 'il-22', 'il-23')
th1 <- c('il-2', 'il-12(p70)', 'il-12(p40)', 'ifng', 'tnfa')

cyt_group_order <- c("Th1", "Th2", "Th17", "Treg","Chemokines", "Other")

# Cytokine name lists for display in heatmaps

# cytokine names capitalized
cyt_names_capital <-  c("IL-2","IL-12p70","IL-12p40","IFN-g","TNF-a",
        "IL-4","IL-5","IL-9","IL-13","IL-25",
        "IL-6","IL-17a","IL-17f","IL-22",
        "IL-10",
        "IL-8","MCP-1","MIP-1a","MIP-1b","RANTES","MIG","IP-10",
        "IL-1a","IL-1b","IL-1Ra","IL-3","IL-15","IL-18","G-CSF","GM-CSF","VEGF-A","sCD40L")

# cytokine names with greek letters
cyt_names_greek <-  c("IL-2","IL-12p70","IL-12p40",paste0("IFN-",greeks("gamma")),paste0("TNF-",greeks("alpha")),
        "IL-4","IL-5","IL-9","IL-13","IL-25",
        "IL-6","IL-17a","IL-17f","IL-22",
        "IL-10",
        "IL-8","MCP-1",paste0("MIP-1",greeks("alpha")),paste0("MIP-1",greeks("beta")),"RANTES","MIG","IP-10",
        paste0("IL-1",greeks("alpha")),paste0("IL-1",greeks("beta")),"IL-1Ra","IL-3","IL-15","IL-18","G-CSF","GM-CSF","VEGF-A","sCD40L")

# cytokine names with greek letters and use of formal name
# Example: MIG = CXCL9
cyt_names_greek_formal <-  c("IL-2","IL-12p70","IL-12p40",paste0("IFN-",greeks("gamma")),paste0("TNF-",greeks("alpha")),
        "IL-4","IL-5","IL-9","IL-13","IL-25",
        "IL-6","IL-17a","IL-17f","IL-22",
        "IL-10",
        "IL-8","CCL2","CCL3","CCL4","CCL5","CXCL9","CXCL10",
        paste0("IL-1",greeks("alpha")),paste0("IL-1",greeks("beta")),"IL-1Ra","IL-3","IL-15","IL-18","G-CSF","GM-CSF","VEGF-A","sCD40L")

```

## Baseline heatmap components

### Transpose and scaling cytokine baseline concentration for heatmap

```{r}

# Transpose tibble
wide_baseline_cyt_heat <- clinical_cytokine_baseline_age.status %>%
  mutate(`ae type` = tolower(`ae type`)) %>%
  mutate(cyt_group = case_when(
    cytokines %in% th1 ~ "Th1",
    cytokines %in% th2 ~ "Th2",
    cytokines %in% th17 ~ "Th17",
    cytokines %in% treg ~ "Treg",
    cytokines %in% chemokines ~ "Chemokines",
    cytokines %in% other ~ "Other",
    TRUE ~ NA_character_)
    ) %>%
  arrange(age65_status) %>%
  select(`pub_id paper`,`cytokines`, `average_baseline_conc`, cyt_group) %>%
  pivot_wider(names_from = `pub_id paper`, values_from = `average_baseline_conc`) %>%
  arrange(match(cytokines, cyt_order))

#extract cytokine col names prior to making dataframe
cyt_names <- as.matrix(wide_baseline_cyt_heat[,1])

#make dataframe for pheatmap
heat_df_wide_cyt_baseline <- as.matrix(
  wide_baseline_cyt_heat[,3:ncol(wide_baseline_cyt_heat)])

#use cytokine names as row names for the pheatmap dataframe
row.names(heat_df_wide_cyt_baseline) <- cyt_names

#two scaling options
heat_df_log2_scale_cyt_baseline = log(heat_df_wide_cyt_baseline, 2)
heat_df_log_scale_cyt_baseline = log(heat_df_wide_cyt_baseline,10)

# scaling using scale function
transpose_heat_df_wide_cyt_baseline <- t(heat_df_wide_cyt_baseline)

scaled_transpose_heat_df_wide_cyt_baseline = scale(transpose_heat_df_wide_cyt_baseline)

heat_df_num_scale_cyt_baseline <- t(scaled_transpose_heat_df_wide_cyt_baseline)

```

### Creation of baseline heatmap label from clinical tibbles

```{r}

pt_id_cyt_baseline <- clinical_cytokine_baseline_age.status %>%
  distinct(`pt id`)

#heatmap labels

baseline_cyt_heat_label <- clinical_tb %>%
  mutate_all(.funs = tolower) %>%
  filter(`pt id` %in% pt_id_cyt_baseline$`pt id`) %>%
  mutate(`type of regimen` = case_when(
        str_detect(`regimen`, "ipilimumab") ~ "CTLA based",
        TRUE ~ "PD1/PDL1")) %>%
  mutate(cancer_group = case_when(`cancer group` == "gi" ~ "gastrointestinal",
                                  `cancer group` == "gu" ~ "genitourinary",
                                  `cancer group` == "skin" ~ "skin",
                                  `cancer group` == "upper aerodigestive" ~ "upper aerodigestive",
                                  TRUE ~ "other")) %>%
  distinct() %>%
  filter(!is.na(`autoimmune history`)) %>%
  mutate(`clinical response` = case_when(str_detect(response, paste(c("^complete response$","^partial response$"), collapse = "|")) ~ "yes",
                         is.na(response) ~ NA_character_,
                         str_detect(response, paste(c("^unknown$","not assessable"), collapse = "|")) ~ NA_character_,
                         TRUE ~ "no")) %>%
  arrange(age65_status) %>%
  select(`pub_id paper`, age65_status,
         `type of regimen`,
         `cancer_group`
         ) 

```

### Creating clinical annotation object for baseline heatmap

```{r}
colors_npg <- pal_npg()(10)

color_from_viridis <- viridis(12, option = "D", alpha = 1)

greater_than_equal <- "\u2265"
no_gr2_irAE <- paste0("No grade ",greater_than_equal,"2 irAE")

cyt_baseline_ann <- data.frame(baseline_cyt_heat_label[,-1])
colnames(cyt_baseline_ann) <- c('Age status',
                   'Type of regimen',
                   'Cancer group')

cyt_baseline_annotate_cols <- list(`Cancer group`= 
                        c(`gastrointestinal` = color_pal_blu.red.yellow[1],
                          `genitourinary` = color_pal_blu.red.yellow[2],
                          `skin` = color_pal_blu.red.yellow[3],
                          `upper aerodigestive` = color_pal_blu.red.yellow[4],
                          `other` = color_pal_blu.red.yellow[5]),
                      `Age status` = c(`yes` = color_pal[2], `no` = color_pal[1]),
                      `Type of regimen` = c(`CTLA based` = "#C96567", `PD1/PDL1` = "#314455")
                      )

cyt_baseline_colAnn <- HeatmapAnnotation(df= cyt_baseline_ann,
                            which = 'col',
                            col = cyt_baseline_annotate_cols,
                            annotation_width = unit(c(1,4), 'cm'),
                            gap = unit(1, 'mm'),
                            annotation_legend_param = list(
                              `Age status` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                 labels = c("Young","Aged")),
                              `Cancer group` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                    labels = c("Gastrointestinal","Genitourinary","Other","Skin","Upper aerodigestive")),
                              `Type of regimen` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                       labels = c("anti-CTLA-4 + anti-PD-1","anti-PD-1 or anti-PD-L1"))
                                                           ))
```

### Set up for arranging the baseline heatmap annotations

```{r}
cyt_group_tb <- wide_baseline_cyt_heat %>%
  select(cytokines, cyt_group)

cyt_group_df <- as.matrix(wide_baseline_cyt_heat$cyt_group)

row.names(cyt_group_df) <- cyt_names

# row splitting for heatmap
# divide by cytokine groups
row_split <- factor(cyt_group_tb$cyt_group, levels = cyt_group_order)

# column splitting for heatmpa
# divide by age status
col_split <- factor(baseline_cyt_heat_label$age65_status, levels = c("no","yes"))

```

## On treatment heatmap components

### Transpose and scaling cytokine on treatment concentration for heatmap

```{r}

#Transpose on.treatment concentration tibble
wide_timepoint1_conc_cyt_heat <- clinical_change_timepoint1_age.status %>%
  mutate(`ae type` = tolower(`ae type`)) %>%
  mutate(cyt_group = case_when(
    cytokines %in% th1 ~ "Th1",
    cytokines %in% th2 ~ "Th2",
    cytokines %in% th17 ~ "Th17",
    cytokines %in% treg ~ "Treg",
    cytokines %in% chemokines ~ "Chemokines",
    cytokines %in% other ~ "Other",
    TRUE ~ NA_character_)
    ) %>%
  arrange(age65_status) %>%
  select(`pub_id paper`, `cytokines`, `concentration`, cyt_group) %>%
  pivot_wider(names_from = `pub_id paper`, values_from = `concentration`) %>%
  arrange(match(cytokines, cyt_order))

#extract cytokine col names prior to making dataframe
cyt_names <- as.matrix(wide_timepoint1_conc_cyt_heat[,1])

#make dataframe for pheatmap
heat_df_wide_conc_cyt_timepoint1 <- as.matrix(
  wide_timepoint1_conc_cyt_heat[,3:ncol(wide_timepoint1_conc_cyt_heat)])

#use cytokine names as row names for the pheatmap dataframe
row.names(heat_df_wide_conc_cyt_timepoint1) <- cyt_names

#two scaling options
heat_df_log2_conc_cyt_timepoint1 = log(heat_df_wide_conc_cyt_timepoint1, 2)
heat_df_log_conc_cyt_timepoint1 = log(heat_df_wide_conc_cyt_timepoint1,10)

# scaling using scale function
transpose_heat_df_wide_conc_cyt_timepoint1 <- t(heat_df_wide_conc_cyt_timepoint1)

scaled.conc_transpose_heat_df_wide_cyt_timepoint1 = scale(transpose_heat_df_wide_conc_cyt_timepoint1)

heat_df_num_scaled.conc_cyt_timepoint1 <- t(scaled.conc_transpose_heat_df_wide_cyt_timepoint1)

```

### Transpose and scaling cytokine on treatment fold change for heatmap

```{r}

#Transpose on.treatment fold change tibble
wide_timepoint1_fold_cyt_heat <- clinical_change_timepoint1_age.status %>%
  mutate(`ae type` = tolower(`ae type`)) %>%
  mutate(cyt_group = case_when(
    cytokines %in% th1 ~ "Th1",
    cytokines %in% th2 ~ "Th2",
    cytokines %in% th17 ~ "Th17",
    cytokines %in% treg ~ "Treg",
    cytokines %in% chemokines ~ "Chemokines",
    cytokines %in% other ~ "Other",
    TRUE ~ NA_character_)
    ) %>%
  arrange(age65_status) %>%
  select(`pub_id paper`, `cytokines`, `fold`, cyt_group) %>%
  pivot_wider(names_from = `pub_id paper`, values_from = `fold`) %>%
  arrange(match(cytokines, cyt_order))

wide_timepoint1_log2.fold_cyt_heat <- clinical_change_timepoint1_age.status %>%
  mutate(`ae type` = tolower(`ae type`)) %>%
  mutate(cyt_group = case_when(
    cytokines %in% th1 ~ "Th1",
    cytokines %in% th2 ~ "Th2",
    cytokines %in% th17 ~ "Th17",
    cytokines %in% treg ~ "Treg",
    cytokines %in% chemokines ~ "Chemokines",
    cytokines %in% other ~ "Other",
    TRUE ~ NA_character_)
    ) %>%
  arrange(age65_status) %>%
  select(`pub_id paper`, `cytokines`, `fold_log2`, cyt_group) %>%
  pivot_wider(names_from = `pub_id paper`, values_from = `fold_log2`) %>%
  arrange(match(cytokines, cyt_order))

#extract cytokine col names prior to making dataframe
cyt_names <- as.matrix(wide_timepoint1_fold_cyt_heat[,1])

#make dataframe for pheatmap
heat_df_wide_fold_cyt_timepoint1 <- as.matrix(
  wide_timepoint1_fold_cyt_heat[,3:ncol(wide_timepoint1_fold_cyt_heat)])

#use cytokine names as row names for the pheatmap dataframe
row.names(heat_df_wide_fold_cyt_timepoint1) <- cyt_names

#two scaling options
heat_df_log2.fold_cyt_timepoint1 = log(heat_df_wide_fold_cyt_timepoint1, 2)
heat_df_log.fold_cyt_timepoint1 = log(heat_df_wide_fold_cyt_timepoint1,10)

```

### Creation of on treatment heatmap labels from clinical tibbles

```{r}
# Pt ID in on treatment cytokine tibble

pt_id_cyt_timepoint1 <- clinical_change_timepoint1_age.status %>%
  distinct(`pt id`)

#heatmap labels

#On treatment cytokine heatmap label
timepoint1_cyt_heat_label <- clinical_tb %>%
  mutate_all(.funs = tolower) %>%
  filter(`pt id` %in% pt_id_cyt_timepoint1$`pt id`) %>%
  mutate(`type of regimen` = case_when(
        str_detect(`regimen`, "ipilimumab") ~ "CTLA based",
        TRUE ~ "PD1/PDL1")) %>%
  mutate(cancer_group = case_when(`cancer group` == "gi" ~ "gastrointestinal",
                                  `cancer group` == "gu" ~ "genitourinary",
                                  `cancer group` == "skin" ~ "skin",
                                  `cancer group` == "upper aerodigestive" ~ "upper aerodigestive",
                                  TRUE ~ "other")) %>%
  distinct() %>%
  filter(!is.na(`autoimmune history`)) %>%
  mutate(`clinical response` = case_when(str_detect(response, paste(c("^complete response$","^partial response$"), collapse = "|")) ~ "yes",
                         is.na(response) ~ NA_character_,
                         str_detect(response, paste(c("^unknown$","not assessable"), collapse = "|")) ~ NA_character_,
                         TRUE ~ "no")) %>%
  arrange(age65_status) %>%
  select(`pub_id paper`, age65_status,
         `type of regimen`,
         `cancer_group`
         ) 

```

### Creating clinical annotation for on treatment heatmaps

```{r}
colors_npg <- pal_npg()(10)

color_from_viridis <- viridis(12, option = "D", alpha = 1)

cyt_timepoint1_ann <- data.frame(timepoint1_cyt_heat_label[,-1])
colnames(cyt_timepoint1_ann) <- c('Age status',
                   'Type of regimen',
                   'Cancer group')

cyt_timepoint1_annotate_cols <- list(`Cancer group`= 
                        c(`gastrointestinal` = color_pal_blu.red.yellow[1],
                          `genitourinary` = color_pal_blu.red.yellow[2],
                          `skin` = color_pal_blu.red.yellow[3],
                          `upper aerodigestive` = color_pal_blu.red.yellow[4],
                          `other` = color_pal_blu.red.yellow[5]),
                      `Age status` = c(`yes` = color_pal[2], `no` = color_pal[1]),
                      `Type of regimen` = c(`CTLA based` = "#C96567", `PD1/PDL1` = "#314455")
                      )

cyt_timepoint1_colAnn <- HeatmapAnnotation(df= cyt_timepoint1_ann,
                            which = 'col',
                            col = cyt_timepoint1_annotate_cols,
                            annotation_width = unit(c(1,4), 'cm'),
                            gap = unit(1, 'mm'),
                            annotation_legend_param = list(
                              `Age status` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                 labels = c("Young","Aged") 
                                                 ),
                              `Cancer group` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                    labels = c("Gastrointestinal","Genitourinary","Other","Skin","Upper aerodigestive")
                                                 ),
                              `Type of regimen` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                       labels = c("anti-CTLA-4 + anti-PD-1","anti-PD-1 or anti-PD-L1")
                                                 )))
```

### Set up for arranging the on treatment heatmaps

```{r}
cyt_group_tb_timepoint1 <- wide_timepoint1_conc_cyt_heat %>%
  select(cytokines, cyt_group)

cyt_group_df_timepoint1 <- as.matrix(wide_timepoint1_conc_cyt_heat$cyt_group)

row.names(cyt_group_df_timepoint1) <- cyt_names

row_split_timepoint1 <- factor(cyt_group_tb$cyt_group, levels = cyt_group_order)

col_split_timepoint1 <- factor(timepoint1_cyt_heat_label$age65_status, levels = c("no","yes"))

```

# Plots

### Baseline cytokine heatmap

```{r,echo = FALSE, warning = FALSE, message = FALSE}

col_fun_baseline = colorRamp2(c(-2,0,2,4,6), c("#2166AC","#F7F7F7","#FDDBC7", "#D6604D", "#B2182B"))

baseline_cyt_scaled_data_heatmap <- Heatmap(heat_df_num_scale_cyt_baseline, 
                                            col = col_fun_baseline, 
                                            heatmap_legend_param = list(title = "Scaled \nconcentration"
                                                                        ),
                                            show_column_names = FALSE,
                                            column_title = " ",
                                            column_title_side = "bottom",
                                            row_title_rot = 0,
                                            row_title_gp = gpar(fontsize = 14), 
                                            row_split = row_split,
                                            column_split = col_split,
                                            row_names_side = "left",
                                            show_row_names = FALSE,
                                            cluster_rows = FALSE, 
                                            cluster_columns = FALSE,
                                            top_annotation=cyt_baseline_colAnn ,
                                            border_gp = gpar(col = "black", lty = 1),
                                            rect_gp = gpar(col = NA, lwd = 2),
                                            )

baseline_cyt_label_heatmap <- Heatmap(cyt_group_df, 
                                      col = colors_cyt_label_heatmap,
                                      show_column_names= FALSE, 
                                      show_row_names = TRUE,
                                      row_split = row_split,
                                      show_heatmap_legend = FALSE,
                                      heatmap_legend_param = list(title = "Cytokine Group"),
                                      row_title_gp = gpar(fontsize = 12))

baseline_cyt_label_heatmap_greek <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = FALSE,
              row_split = row_split,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12)) +
  rowAnnotation(labels = anno_text(cyt_names_greek))

baseline_cyt_label_heatmap_capital <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = FALSE,
              row_split = row_split,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12)) +
  rowAnnotation(labels = anno_text(cyt_names_capital))

baseline_cyt_label_heatmap_greek_formal <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = FALSE,
              row_split = row_split,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12)) +
  rowAnnotation(labels = anno_text(cyt_names_greek_formal))

baseline_scaled_total_heatmap <- baseline_cyt_label_heatmap + baseline_cyt_scaled_data_heatmap

baseline_scaled_total_heatmap_greek <- baseline_cyt_label_heatmap_greek + baseline_cyt_scaled_data_heatmap

baseline_scaled_total_heatmap_capital <- baseline_cyt_label_heatmap_capital + baseline_cyt_scaled_data_heatmap

baseline_scaled_total_heatmap_greek_formal <- baseline_cyt_label_heatmap_greek_formal + baseline_cyt_scaled_data_heatmap

draw(baseline_scaled_total_heatmap_greek_formal, auto_adjust = FALSE)
```

### On treatment heatmap (scaled concentration or zscore)

```{r,echo = FALSE, warning = FALSE, message = FALSE}
col_fun_timepoint1 = colorRamp2(c(-2,0,2,4,6), c("#2166AC","#F7F7F7","#FDDBC7", "#D6604D", "#B2182B"))

timepoint1_cyt_scaled.conc_data_heatmap <- Heatmap(heat_df_num_scaled.conc_cyt_timepoint1, 
        col = col_fun_timepoint1,
        heatmap_legend_param = list(title = "Scaled \nconcentration"),
        show_column_names = FALSE,
        column_title = " ",
        column_title_side = "bottom",
        row_title_rot = 0,
        row_title_gp = gpar(fontsize = 14), 
        row_split = row_split_timepoint1,
        column_split = col_split_timepoint1,
        row_names_side = "left",
        show_row_names = FALSE,
        cluster_rows = FALSE, 
        cluster_columns = FALSE,
        top_annotation=cyt_timepoint1_colAnn ,
        border_gp = gpar(col = "black", lty = 1),
        rect_gp = gpar(col = NA, lwd = 2),
        )

timepoint1_cyt_label_heatmap <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = TRUE,
              row_split = row_split_timepoint1,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12))

timepoint1_cyt_label_heatmap_greek <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = FALSE,
              row_split = row_split_timepoint1,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12)) +
  rowAnnotation(labels = anno_text(cyt_names_greek))

timepoint1_cyt_label_heatmap_capital <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = FALSE,
              row_split = row_split,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12)) +
  rowAnnotation(labels = anno_text(cyt_names_capital))

timepoint1_cyt_label_heatmap_greek_formal <- Heatmap(cyt_group_df, 
              col = colors_cyt_label_heatmap,
              show_column_names= FALSE, 
              show_row_names = FALSE,
              row_split = row_split_timepoint1,
              show_heatmap_legend = FALSE,
              heatmap_legend_param = list(title = "Cytokine Group"),
              row_title_gp = gpar(fontsize = 12)) +
  rowAnnotation(labels = anno_text(cyt_names_greek_formal))

timepoint1_scaled.conc_total_heatmap <- timepoint1_cyt_label_heatmap + timepoint1_cyt_scaled.conc_data_heatmap

timepoint1_scaled.conc_total_heatmap_greek <- timepoint1_cyt_label_heatmap_greek + timepoint1_cyt_scaled.conc_data_heatmap

timepoint1_scaled.conc_total_heatmap_capital <- timepoint1_cyt_label_heatmap_capital + timepoint1_cyt_scaled.conc_data_heatmap

timepoint1_scaled.conc_total_heatmap_greek_formal <- timepoint1_cyt_label_heatmap_greek_formal + timepoint1_cyt_scaled.conc_data_heatmap

draw(timepoint1_scaled.conc_total_heatmap_greek_formal,auto_adjust = FALSE)
```

### On treatment cytokine heatmap (log 2 fold change)

```{r,echo = FALSE, warning = FALSE, message = FALSE}

timepoint1_cyt_fold.log2_data_heatmap <- Heatmap(heat_df_log2.fold_cyt_timepoint1, 
        col = colorRampPalette(c("navy","white","firebrick3"))(50), 
        heatmap_legend_param = list(title = expression(Log[2] ~ "fold")
                                    ),
        show_column_names = FALSE,
        column_title = " ",
        column_title_side = "bottom",
        row_title_rot = 0,
        row_title_gp = gpar(fontsize = 14), 
        row_split = row_split_timepoint1,
        column_split = col_split_timepoint1,
        row_names_side = "left",
        show_row_names = FALSE,
        cluster_rows = FALSE, 
        cluster_columns = FALSE,
        top_annotation=cyt_timepoint1_colAnn ,
        border_gp = gpar(col = "black", lty = 1),
        rect_gp = gpar(col = NA, lwd = 2),
        )

timepoint1_fold.log2_total_heatmap <- timepoint1_cyt_label_heatmap + timepoint1_cyt_fold.log2_data_heatmap

#draw(timepoint1_fold.log2_total_heatmap,auto_adjust = FALSE)

timepoint1_fold.log2_total_heatmap_greek <- timepoint1_cyt_label_heatmap_greek + timepoint1_cyt_fold.log2_data_heatmap

#draw(timepoint1_fold.log2_total_heatmap_greek,auto_adjust = FALSE)

timepoint1_fold.log2_total_heatmap_capital <- timepoint1_cyt_label_heatmap_capital + timepoint1_cyt_fold.log2_data_heatmap

#draw(timepoint1_fold.log2_total_heatmap_capital,auto_adjust = FALSE)

timepoint1_fold.log2_total_heatmap_greek_formal <- timepoint1_cyt_label_heatmap_greek_formal + timepoint1_cyt_fold.log2_data_heatmap

draw(timepoint1_fold.log2_total_heatmap_greek_formal,auto_adjust = FALSE)
```

# Plot saving

## Baseline scaled concentration (Ext Figure 2A)

### Cairo pdf save

```{r}
# Open cairo_pdf device
cairo_pdf(file = paste0(output.path.pdf, "/", Sys.Date(), "_", "baseline_scaled.conc_cyt_aging_heatmap_greek_formal.pdf"),
     height = 8, width = 12)

# print heatmap to device
draw(baseline_scaled_total_heatmap_greek_formal,auto_adjust = FALSE)

# Close device
dev.off()

```

### TIFF save

```{r}
# Open TIFF device
tiff(filename = paste0(output.path.tiff, "/", Sys.Date(), "_", "baseline_scaled.conc_cyt_aging_heatmap_greek_formal.tiff"),
     height = 8, width = 12, units = "in", res = 600)

# Print the plot to the TIFF device
draw(baseline_scaled_total_heatmap_greek_formal,auto_adjust = FALSE)

# Close the TIFF device
dev.off()

```

## On treatment scaled concentration (Ext Figure 2B)

### Cairo pdf save

```{r}
# Open cairo_pdf device
cairo_pdf(file = paste0(output.path.pdf, "/", Sys.Date(), "_", "timepoint1_scaled.conc_cyt_aging_heatmap_greek_formal.pdf"),
     height = 8, width = 12)

# print heatmap to device
draw(timepoint1_scaled.conc_total_heatmap_greek_formal, #annotation_legend_side = "bottom", 
     auto_adjust = FALSE)

# Close device
dev.off()

```

### TIFF save

```{r}
# Open TIFF device
tiff(filename = paste0(output.path.tiff, "/", Sys.Date(), "_", "timepoint1_scaled.conc_cyt_aging_heatmap_greek_formal.tiff"),
     height = 8, width = 12, units = "in", res = 600)

# Print the plot to the TIFF device
draw(timepoint1_scaled.conc_total_heatmap_greek_formal, #annotation_legend_side = "bottom", 
     auto_adjust = FALSE)

# Close the TIFF device
dev.off()

```

## On treatment log2 fold (Figure 2B)

### Cairo pdf save

```{r}
# Open cairo_pdf device
cairo_pdf(file = paste0(output.path.pdf, "/", Sys.Date(), "_", "timepoint1_log2fold_cyt_aging_heatmap_greek_formal.pdf"),
     height = 8, width = 12)

# print heatmap to device
draw(timepoint1_fold.log2_total_heatmap_greek_formal,auto_adjust = FALSE)

# Close device
dev.off()

```

### TIFF save

```{r}
# Open TIFF device
tiff(filename = paste0(output.path.tiff, "/", Sys.Date(), "_", "timepoint1_log2fold_cyt_aging_heatmap_greek.tiff"),
     height = 8, width = 18, units = "in", res = 600)

# Print the plot to the TIFF device
draw(timepoint1_fold.log2_total_heatmap_greek_formal,auto_adjust = FALSE)

# Close the TIFF device
dev.off()

```

### Supporting data

#### Supporting data generation

```{r}
# baseline
transpose_heat_df_num_scale_cyt_baseline <- heat_df_num_scale_cyt_baseline %>%
  t(.)

colnames(transpose_heat_df_num_scale_cyt_baseline) <- cyt_names_greek_formal

transpose_baseline_num_scale_pt_order <- rownames(transpose_heat_df_num_scale_cyt_baseline)

baseline_cyt_scaled.conc_supporting.data <-transpose_heat_df_num_scale_cyt_baseline %>%
  cbind(transpose_baseline_num_scale_pt_order,.) %>%
  as_tibble() %>%
  rename(`pub_id paper` = "transpose_baseline_num_scale_pt_order") %>%
  mutate(`pub_id paper` = tolower(`pub_id paper`)) %>%
  left_join(baseline_cyt_heat_label, by = "pub_id paper") %>%
  mutate(`pub_id paper` = toupper(`pub_id paper`)) %>%
  mutate(cancer_group = str_to_title(cancer_group),
         age65_status = str_to_title(age65_status)) %>%
  relocate(age65_status:cancer_group,.after = `pub_id paper`) %>%
  rename(`Cancer group` = "cancer_group",
         `Aged status` = "age65_status",
         `Type of regimen` = "type of regimen",
         `Subject ID` = "pub_id paper")

# on treatment scaled concentration

on.treatment_num_scale_pt_order <- rownames(scaled.conc_transpose_heat_df_wide_cyt_timepoint1)

on.treatment_cyt_scaled.conc_supporting.data <-scaled.conc_transpose_heat_df_wide_cyt_timepoint1 %>%
  cbind(on.treatment_num_scale_pt_order,.) %>%
  as_tibble() %>%
  rename(`pub_id paper` = "on.treatment_num_scale_pt_order") %>%
  mutate(`pub_id paper` = tolower(`pub_id paper`)) %>%
  left_join(timepoint1_cyt_heat_label, by = "pub_id paper") %>%
  mutate(`pub_id paper` = toupper(`pub_id paper`)) %>%
  mutate(cancer_group = str_to_title(cancer_group),
         age65_status = str_to_title(age65_status)) %>%
  relocate(age65_status:cancer_group,.after = `pub_id paper`) %>%
  rename(`Cancer group` = "cancer_group",
         `Aged status` = "age65_status",
         `Type of regimen` = "type of regimen",
         `Subject ID` = "pub_id paper")

# On treatment log2 fold

transpose_heat_df_log2.fold_cyt_timepoint1 <- heat_df_log2.fold_cyt_timepoint1 %>%
  t(.)

colnames(transpose_heat_df_log2.fold_cyt_timepoint1) <- cyt_names_greek_formal

transpose_timepoint1_log2.fold_pt_order <- rownames(transpose_heat_df_log2.fold_cyt_timepoint1)

on.treatment_cyt_log2fold_supporting.data <- transpose_heat_df_log2.fold_cyt_timepoint1 %>%
  cbind(transpose_timepoint1_log2.fold_pt_order,.) %>%
  as_tibble() %>%
  rename(`pub_id paper` = "transpose_timepoint1_log2.fold_pt_order") %>%
  mutate(`pub_id paper` = tolower(`pub_id paper`)) %>%
  left_join(timepoint1_cyt_heat_label, by = "pub_id paper") %>%
  mutate(`pub_id paper` = toupper(`pub_id paper`)) %>%
  mutate(cancer_group = str_to_title(cancer_group),
         age65_status = str_to_title(age65_status)) %>%
  relocate(age65_status:cancer_group,.after = `pub_id paper`) %>%
  rename(`Cancer group` = "cancer_group",
         `Aged status` = "age65_status",
         `Type of regimen` = "type of regimen",
         `Subject ID` = "pub_id paper")
```

### Exporting 

```{r}

write.xlsx(baseline_cyt_scaled.conc_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "baseline_scaled.conc_aging_heatmap_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)

write.xlsx(on.treatment_cyt_scaled.conc_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "on.treatment_scaled.conc_aging_heatmap_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)

write.xlsx(on.treatment_cyt_log2fold_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "timepoint1_log2fold_cyt_aging_heatmap_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)

```