---
title: "CyTOF Marker Heatmap"
output: html_document
date: "2024-05-23"
---
# Set up

### Libraries

```{r include = FALSE}
# Libraries
library(here)
library(tidyverse)
library(ggplot2)
library(pheatmap)
library(viridis)
library(ggpubr)
library(gridExtra)
library(grid)
library(ggbreak) 
library(ggsci)
library(ComplexHeatmap)
library(khroma)
library(greekLetters)
library(Cairo)
library(RColorBrewer)
library(readxl)
library(circlize)
library(openxlsx)
library(NatParksPalettes)

```

### Directory pathing

```{r}
here::i_am("code/02.7_aging_cytof.marker_heatmaps.Rmd")

output.path.rds <- here("output/RDS files")
if (!dir.exists(here(output.path.rds))) dir.create(here(output.path.rds))

output.path.excel <- here("output/Excel files")
if (!dir.exists(here(output.path.excel))) dir.create(here(output.path.excel))

output.path.figure_files <- here("output/Figure files")
if (!dir.exists(here(output.path.figure_files))) dir.create(here(output.path.figure_files))

output.path.tiff <- here("output/Figure files/Tiff files")
if (!dir.exists(here(output.path.tiff))) dir.create(here(output.path.tiff))

output.path.svg <- here("output/Figure files/SVG files")
if (!dir.exists(here(output.path.svg))) dir.create(here(output.path.svg))

output.path.pdf <- here("output/Figure files/PDF files")
if (!dir.exists(here(output.path.pdf))) dir.create(here(output.path.pdf))

```

### Import files

```{r}

clinical<- readRDS(here("output/RDS files/clinical.rds"))

clinical$orr.age.status <- factor(clinical$orr.age.status, 
                                                levels = c("RY", "NRY",
                                                           "RO", "NRO"))

file_path_v1 <- here("data/CyTOF/V1_Functional_Marker_Expression.xlsx")
cytof.marker.tb_import_v1 <- read_excel(file_path_v1)

```

### Data cleaning

```{r}

# version 1 clustering
cytof.marker.tb_v1 <- cytof.marker.tb_import_v1 %>%
  filter(!sample_id == "15_0") %>%
  mutate(sample_id = ifelse(sample_id == "15_0_1", "15_0", sample_id)) %>%
  mutate(`pt id` = sub("^(\\d+)_.*", "\\1",`sample_id`)) %>%
  mutate(`pt id` = as.numeric(`pt id`)) %>%
  mutate(`timepoint` = sub(".*_", "\\1",`sample_id`)) %>%
  mutate(`timepoint` = as.numeric(`timepoint`)) %>%
  relocate(`pt id`, .before = cluster) %>%
  rename(immune_cell = "cluster", marker = "variable", mmi = "value") 

change_cytof.marker.tb_v1 <- cytof.marker.tb_v1 %>%
  group_by(`pt id`,`immune_cell`,`marker`) %>%
  arrange(`pt id`,`immune_cell`,`marker`,`timepoint`) %>%
  mutate(baseline_mmi = mmi[row_number() ==1],
         delta = mmi - mmi[row_number() == 1],
         relative_change = (mmi - mmi[row_number() == 1])/mmi[row_number() == 1],
         fold = mmi/mmi[row_number() == 1],
         fold_log2 = log2(fold)) %>%
  filter(!`timepoint` == 0) %>%
  relocate(baseline_mmi:fold, .after = mmi) %>%
  ungroup()

```

### Combining files

```{r}

# v1 clustering

clinical_cytof.marker_tb_v1 <- clinical %>%
  left_join(cytof.marker.tb_v1, by = "pt id") %>%
  mutate(`blood timepoint` = case_when(timepoint == 0 ~ "baseline",
                                       timepoint == 1 ~ "on treatment",
                                       TRUE ~ NA_character_))

clinical_cytof.change_tb_v1 <- clinical %>%
  left_join(change_cytof.marker.tb_v1, by = "pt id") %>%
  mutate(`blood timepoint` = case_when(timepoint == 0 ~ "baseline",
                                       timepoint == 1 ~ "on treatment",
                                       TRUE ~ NA_character_))

```

### Cytof cluster lists (grouped)

```{r}

cytof_cluster_grouped_list <- unique(clinical_cytof.marker_tb_v1$immune_cell)
cytof_cluster_grouped_list_rmv_UA <- cytof_cluster_grouped_list[-16]

```

### Baseline and on treatment tibbles (grouped clustering version 1)

```{r}
# baseline tibble
clinical_cytof.marker_baseline_v1 <- clinical_cytof.marker_tb_v1 %>%
  filter(timepoint == 0) %>%
  group_by(immune_cell, marker, age65_status) %>%
  mutate(mean_age.status_immune.group_marker = mean(mmi)) %>%
  group_by(immune_cell, marker) %>%
  mutate(mean_immune.group_marker = mean(mmi),
         sd_immune.group_marker = sd(mmi)) %>%
  ungroup() %>%
  mutate(zscore_pt = (mmi-mean_immune.group_marker)/sd_immune.group_marker) %>%
  group_by(immune_cell, marker, age65_status) %>% 
  mutate(mean_zscore = mean(zscore_pt))

# first timepoint tibble
clinical_cytof.marker_on.treatment_v1 <- clinical_cytof.marker_tb_v1 %>%
  filter(timepoint == 1) %>%
  group_by(immune_cell, marker, age65_status) %>%
  mutate(mean_age.status_immune.group_marker = mean(mmi)) %>%
  group_by(immune_cell, marker) %>%
  mutate(mean_immune.group_marker = mean(mmi),
         sd_immune.group_marker = sd(mmi)) %>%
  ungroup() %>%
  mutate(zscore_pt = (mmi-mean_immune.group_marker)/sd_immune.group_marker) %>%
  group_by(immune_cell, marker, age65_status) %>% 
  mutate(mean_zscore = mean(zscore_pt))

# baseline and first timepoint tibble
clinical_cytof.marker_baseline.ontreatment_v1 <- clinical_cytof.marker_tb_v1 %>%
  filter(timepoint %in% c(0,1)) %>%
  group_by(immune_cell, marker, age65_status) %>%
  mutate(mean_age.status_immune.group_marker = mean(mmi)) %>%
  group_by(timepoint,immune_cell, marker) %>%
  mutate(mean_immune.group_marker = mean(mmi),
         sd_immune.group_marker = sd(mmi)) %>%
  ungroup() %>%
  mutate(zscore_pt = (mmi-mean_immune.group_marker)/sd_immune.group_marker) %>%
  group_by(timepoint,immune_cell, marker, age65_status) %>% 
  mutate(mean_zscore = mean(zscore_pt))

```

### Special characters for graphs

```{r}
# For >= symbol in strings
greater_than_equal <- "\u2265"
less_than <- "\u003c"

age_greater.equal_65 <- paste0("Age ",greater_than_equal,"65")

age_less_65 <- paste0("Age ",less_than,"65")

```

### Color palettes

```{r}
colors_npg <- pal_npg()(10)

colors_anno_label_heatmap <- c("#544667","#C37672","#F49835","#4AB8A7","#284B96","#878787")

color_pal_custom <- c(#"#84D7E199","#8A419899","#FF95A899",
  #"#374E56FF","#DF8F46FF","#00A1D5FF",
  #"#003049","#0a9396","#94d2bd","#ffb703","#f77100","#d62826",
  #"#000181",
  #"#1974D3",
  "#219ebc","#023047","#ffb703",
  "#d1495b","#edae49","#00798c",
  "#001219",
  #"#005f73",
  "#0a9396",
  #"#e9d8a6",
  "#ee9b00",
  "#53ccec",
  #"#F2C858",
  "#F54952","#3A0751",
  "#422680","#280659","#FC666F"
  )

```

### Annotation label for clusters individually

```{r}

total_anno_tb_individual  <- tibble(`Marker status` = c("memory","memory","memory",
                                                           "cytotoxicity","cytotoxicity","cytotoxicity",
                                    rep("exhaustion",6)))

total_anno_cols_individual <- list(`Marker status` = 
                        c(`memory` = color_pal_custom[1],
                          `cytotoxicity` = color_pal_custom[2],
                          `exhaustion` = color_pal_custom[3])
                      )

total_anno_individual <- HeatmapAnnotation(df= total_anno_tb_individual,
                            which = 'col',
                            col = total_anno_cols_individual,
                            annotation_width = unit(c(1,4), 'cm'),
                            show_annotation_name = FALSE,
                            #annotation_name_side = "right",
                            gap = unit(1, 'mm'),
                            annotation_legend_param = list(
                              `Marker status` = list(title_gp = gpar(fontface = "plain"), labels_gp = gpar(fontsize = 12),
                                                    labels = c("Effector function/proliferation",
                                                               "Exhaustion","Memory"),
                                                    direction = "horizontal"
                                                 )),
                              border = TRUE
                              ) 

```

### Markers to use

```{r}
select_markers <- c("CCR7","CD45RA","CD45RO",
                    "GZMB","KI67","TBET",
                    "CTLA4","LAG3","PD1","PDL1",
                    "TIGIT", "TIM3")
```

# Plots

### For loop for heatmaps (Figure 5 and Ext Data Figure 4)

```{r}

col_fun_fixed = colorRamp2(seq(from = -0.5, to = 0.5, by = 0.25), c("#023047","#219ebc","white", "#db504a", "#9b2226"))

# generate empty lists
baseline_plot_list <- list()
on.treatment_plot_list <- list()

#extract immune group col names prior to making dataframe

age65.group_names_v1 <- c("Young","Aged")

for (i in 1:length(cytof_cluster_grouped_list_rmv_UA)) {
  
  baseline_wide_cytof.marker_df <- clinical_cytof.marker_baseline.ontreatment_v1 %>%
    filter(timepoint == 0) %>%
    distinct(immune_cell, marker, age65_status, .keep_all = TRUE) %>%
    select(immune_cell,age65_status,mean_zscore,`marker`, mean_age.status_immune.group_marker, mean_immune.group_marker, sd_immune.group_marker) %>%
    group_by(marker) %>%
    filter(immune_cell == paste(cytof_cluster_grouped_list_rmv_UA[i])) %>%
    filter(marker %in% select_markers) %>%
    arrange(match(marker, select_markers)) %>%
    ungroup() %>%
    select(age65_status,`marker`, mean_zscore) %>%
    mutate(age65_status = ifelse(age65_status == "yes", "Aged", "Young")) %>%
    arrange(desc(age65_status)) %>%
    pivot_wider(names_from = marker, values_from = mean_zscore) 
  
  
  baseline_heat_df <- as.matrix(
  baseline_wide_cytof.marker_df[,2:ncol(baseline_wide_cytof.marker_df)])
  
  row.names(baseline_heat_df) <- age65.group_names_v1
  
  on.treatment_wide_cytof.marker_df <- clinical_cytof.marker_baseline.ontreatment_v1 %>%
    filter(timepoint == 1) %>%
    distinct(immune_cell, marker, age65_status, .keep_all = TRUE) %>%
    select(immune_cell,age65_status,mean_zscore,`marker`, mean_age.status_immune.group_marker, mean_immune.group_marker, sd_immune.group_marker) %>%
    group_by(marker) %>%
    filter(immune_cell == paste(cytof_cluster_grouped_list_rmv_UA[i])) %>%
    filter(marker %in% select_markers) %>%
    arrange(match(marker, select_markers)) %>%
    ungroup() %>%
    select(age65_status,`marker`, mean_zscore) %>%
    mutate(age65_status = ifelse(age65_status == "yes", "Aged", "Young")) %>%
    arrange(desc(age65_status)) %>%
    pivot_wider(names_from = marker, values_from = mean_zscore) 
  
  max_on.treatment_normalized.mmi <- round(max(on.treatment_wide_cytof.marker_df[,-1]),1)
  
  on.treatment_heat_df <- as.matrix(
  on.treatment_wide_cytof.marker_df[,2:ncol(on.treatment_wide_cytof.marker_df)])
  
  row.names(on.treatment_heat_df) <- age65.group_names_v1
  
  total_heat_df <- rbind(baseline_heat_df,on.treatment_heat_df)
  
  row_split <- factor(c("Baseline","Baseline","On treatment","On treatment"), levels = c("Baseline","On treatment"))
  
  total_heatmap <- Heatmap(total_heat_df,
                                  col = col_fun_fixed,
                                  heatmap_legend_param = list(title = "Average \nscaled MMI",
                                                              title_gp = gpar(fontface = "plain"),
                                                              direction = "horizontal"
                                                              ),
                                  column_title = paste(cytof_cluster_grouped_list_rmv_UA[i]),
                                  column_title_side = "top",
                                  column_title_gp = gpar(fontsize = 14),
                                  column_names_rot = 45,
                                  column_names_gp = gpar(fontsize = 10),
                                  row_title_rot = 0,
                                  row_title_gp = gpar(fontsize = 14), 
                                  row_split = row_split,
                                  row_names_side = "right",
                                  show_row_names = TRUE,
                                  cluster_rows = FALSE, 
                                  cluster_columns = FALSE,
                                  top_annotation= total_anno_individual,
                                  border_gp = gpar(col = "black", lty = 1, lwd = 2),
                                  rect_gp = gpar(col = "black", lwd = 2),
                                  )
  
  print(paste(cytof_cluster_grouped_list_rmv_UA[i]))
  
  cairo_pdf(file = paste0(output.path.pdf, "/", Sys.Date(), "_", paste(cytof_cluster_grouped_list_rmv_UA[i]), "_total__marker_version1_heatmap.pdf"),
     height = 5, width = 8)
  
  draw(total_heatmap, row_gap = unit(3, "mm"), merge_legend = TRUE,ht_gap = unit(7, "mm"),annotation_legend_side = "right", heatmap_legend_side = "bottom",
     auto_adjust = TRUE)
  
  dev.off()
}
```

### Generation of supporting data file

```{r}

cytof.marker_heatmap_supporting.data <- clinical_cytof.marker_baseline.ontreatment_v1 %>%
  filter(marker %in% select_markers) %>%
  select(timepoint,age65_status, immune_cell, marker, mean_zscore) %>%
  mutate(age65_status = str_to_title(age65_status),
         timepoint = ifelse(timepoint == "0", "Baseline", "On treatment")) %>%
  ungroup() %>%
  distinct(timepoint,age65_status, immune_cell, marker, .keep_all = TRUE) %>%
  arrange(match(marker, select_markers)) %>%
  pivot_wider(names_from = marker, values_from = mean_zscore) %>%
  arrange(timepoint, immune_cell) %>%
  rename(`Immune cluster` = "immune_cell",
         `Aged status` = "age65_status",
         `Timepoint` = "timepoint")


write.xlsx(cytof.marker_heatmap_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "cytof.marker_avg.zscore_grouped.cluster_heatmpas_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)
```


