---
title: "CyTOF Marker statistical analysis"
output: html_document
date: "2024-05-23"
---

# Set up

### Libraries

```{r}
library(here)
library(tidyverse)
library(survminer)
library(survival)
library(grid)
library(openxlsx)
library(readxl)
library(ggbreak) 
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(svglite)
library(rstatix)
library(viridis)
library(ggsci)
library(NatParksPalettes)

```

### Directory pathing

```{r}
here::i_am("code/02.8_aging_cytof.marker_analysis.Rmd")

output.path.rds <- here("output/RDS files")
if (!dir.exists(here(output.path.rds))) dir.create(here(output.path.rds))

output.path.excel <- here("output/Excel files")
if (!dir.exists(here(output.path.excel))) dir.create(here(output.path.excel))

output.path.figure_files <- here("output/Figure files")
if (!dir.exists(here(output.path.figure_files))) dir.create(here(output.path.figure_files))

output.path.tiff <- here("output/Figure files/Tiff files")
if (!dir.exists(here(output.path.tiff))) dir.create(here(output.path.tiff))

output.path.svg <- here("output/Figure files/SVG files")
if (!dir.exists(here(output.path.svg))) dir.create(here(output.path.svg))

output.path.pdf <- here("output/Figure files/PDF files")
if (!dir.exists(here(output.path.pdf))) dir.create(here(output.path.pdf))
```

### Source functions

```{r}
source(here("code/functions/function_boxplot.R"))
```

### Import files

```{r}

clinical <- readRDS(here("output/RDS files/clinical.rds"))

clinical$orr.age.status <- factor(clinical$orr.age.status, 
                                                levels = c("RO", "NRO",
                                                           "RY", "NRY"))

file_path_v1 <- here("data/CyTOF/V1_Functional_Marker_Expression.xlsx")
cytof.marker.tb_import_v1 <- read_excel(file_path_v1)

```

### Special characters for graphs

```{r}
# For >= symbol in strings
greater_than_equal <- "\u2265"
less_than <- "\u003c"

age_greater.equal_65 <- paste0("Age ",greater_than_equal,"65")
age_greater.equal_65 <- "Aged"

age_less_65 <- paste0("Age ",less_than,"65")
age_less_65 <- "Young"

```

### Color palettes

```{r}
# Set up colors

color_pal <- c("#15607a","#3cc4ad")
```

### Markers

```{r}
select_markers <- c("CCR7","CD45RA","CD45RO","CTLA4",
                    "GZMB","Ki67",
                    "LAG3","PD1","PDL1",
                    "TBET", "TIGIT", "TIM3")
```

# Data cleaning

### String and mmi change calculations

```{r}

# version 1 clustering
cytof.marker.tb_v1 <- cytof.marker.tb_import_v1 %>%
  filter(!sample_id == "15_0") %>%
  mutate(sample_id = ifelse(sample_id == "15_0_1", "15_0", sample_id)) %>%
  mutate(`pt id` = sub("^(\\d+)_.*", "\\1",`sample_id`)) %>%
  mutate(`pt id` = as.numeric(`pt id`)) %>%
  mutate(`timepoint` = sub(".*_", "\\1",`sample_id`)) %>%
  mutate(`timepoint` = as.numeric(`timepoint`)) %>%
  relocate(`pt id`, .before = cluster) %>%
  rename(immune_cell = "cluster", marker = "variable", mmi = "value") 

change_cytof.marker.tb_v1 <- cytof.marker.tb_v1 %>%
  group_by(`pt id`,`immune_cell`,`marker`) %>%
  arrange(`pt id`,`immune_cell`,`marker`,`timepoint`) %>%
  mutate(baseline_mmi = mmi[row_number() ==1],
         delta = mmi - mmi[row_number() == 1],
         relative_change = (mmi - mmi[row_number() == 1])/mmi[row_number() == 1],
         fold = mmi/mmi[row_number() == 1],
         fold_log2 = log2(fold)) %>%
  filter(!`timepoint` == 0) %>%
  relocate(baseline_mmi:fold, .after = mmi) %>%
  ungroup()

```

### Combining files

```{r}

# v1 clustering

clinical_cytof.marker_tb_v1 <- clinical %>%
  left_join(cytof.marker.tb_v1, by = "pt id") %>%
  mutate(`blood timepoint` = case_when(timepoint == 0 ~ "baseline",
                                       timepoint == 1 ~ "on treatment",
                                       TRUE ~ NA_character_))

clinical_cytof.change_tb_v1 <- clinical %>%
  left_join(change_cytof.marker.tb_v1, by = "pt id") %>%
  mutate(`blood timepoint` = case_when(timepoint == 0 ~ "baseline",
                                       timepoint == 1 ~ "on treatment",
                                       TRUE ~ NA_character_))

```

### Baseline and first on treatment

```{r}


clinical_cytof.marker_baseline_v1 <- clinical_cytof.marker_tb_v1 %>%
  filter(timepoint == 0)

clinical_cytof.marker_on.treatment_v1 <- clinical_cytof.marker_tb_v1 %>%
  filter(timepoint == 1)

```

### Cytof baseline and on treatment number of patients

```{r}
clinical_cytof.marker_baseline_v1 %>%
  ungroup() %>%
  distinct(`pt id`, .keep_all = TRUE) %>%
  group_by(age65_status) %>%
  summarise(count = n())

clinical_cytof.marker_on.treatment_v1 %>%
  ungroup() %>%
  distinct(`pt id`, .keep_all = TRUE) %>%
  group_by(age65_status) %>%
  summarise(count = n())
```

### Cytof cluster lists (grouped)

```{r}

cytof_cluster_grouped_list <- unique(clinical_cytof.marker_tb_v1$immune_cell)
cytof_cluster_grouped_list_rmv_UA <- cytof_cluster_grouped_list[-16]

```

# Statistical analysis

### Wilcoxon for baseline

```{r}

Wilcoxon.baseline.marker_mmi_age65status_v1 <- clinical_cytof.marker_baseline_v1  %>%
  filter(!is.na(mmi)) %>%
  filter(marker %in% select_markers) %>%
  group_by(marker,immune_cell) %>%
  transmute(mean_age65.older = mean(mmi[age65_status == "yes"]),
         mean_age65.younger = mean(mmi[age65_status == "no"]),
         std_age65.older = sd(mmi[age65_status == "yes"]),
         std_age65.younger = sd(mmi[age65_status == "no"]),
         pval = wilcox.test(mmi ~ age65_status, paired = FALSE)$p.value) %>%
  distinct(marker,immune_cell, .keep_all = TRUE) %>%
  ungroup() %>%
  adjust_pvalue(.,method="fdr") %>%
  arrange(pval) %>%
  add_significance() %>%
  filter(pval <0.05) 

```
### Wilcoxon for on.treatment

```{r}

Wilcoxon.on.treatment.marker_mmi_age65status_v1 <- clinical_cytof.marker_on.treatment_v1   %>%
  filter(!is.na(mmi)) %>%
  filter(marker %in% select_markers) %>%
  group_by(marker,immune_cell) %>%
  transmute(mean_age65.older = mean(mmi[age65_status == "yes"]),
         mean_age65.younger = mean(mmi[age65_status == "no"]),
         std_age65.older = sd(mmi[age65_status == "yes"]),
         std_age65.younger = sd(mmi[age65_status == "no"]),
         pval = wilcox.test(mmi ~ age65_status, paired = FALSE)$p.value) %>%
  distinct(marker,immune_cell, .keep_all = TRUE) %>%
  ungroup() %>%
  adjust_pvalue(.,method="fdr") %>%
  arrange(pval) %>%
  add_significance() %>%
  filter(pval <0.05) 

```

# Plots (Figure 5)

### Boxplot generation for baseline/on.treatment (by immune cluster)
Faceted plot of all significant markers for that cluster of interest

```{r}
sig.markers_by.cluster <- c(Wilcoxon.baseline.marker_mmi_age65status_v1$marker,Wilcoxon.on.treatment.marker_mmi_age65status_v1$marker) %>%
  unique(.)

baseline_plot_list <- list()

on.treatment_plot_list <- list()

for (i in 1:length(cytof_cluster_grouped_list_rmv_UA)) {
  
  sig_wilcoxon_baseline_per.cluster <- Wilcoxon.baseline.marker_mmi_age65status_v1 %>%
    filter(immune_cell == paste(cytof_cluster_grouped_list_rmv_UA[i]))
  
  sig_wilcoxon_on.treatment_per.cluster <- Wilcoxon.on.treatment.marker_mmi_age65status_v1 %>%
    filter(immune_cell == paste(cytof_cluster_grouped_list_rmv_UA[i]))
  
  sig.markers_by.cluster <- c(sig_wilcoxon_baseline_per.cluster$marker,sig_wilcoxon_on.treatment_per.cluster$marker) %>%
  unique(.) 
  
  dataset_dummy_1 <- clinical_cytof.marker_baseline_v1 %>%
    filter(immune_cell == paste(cytof_cluster_grouped_list_rmv_UA[i])) %>%
    filter(marker %in% sig.markers_by.cluster)
  
  #dataset_dummy_1$marker <- factor(dataset_dummy_1$marker, levels =c("CCR7","CD45RA","CD45RO",
   #                 "GZMB","KI67","TBET",
    #                "CTLA4","LAG3","PD1","PDL1",
     #               "TIGIT", "TIM3"))
  
  dataset_dummy_2 <- clinical_cytof.marker_on.treatment_v1 %>%
    filter(immune_cell == paste(cytof_cluster_grouped_list_rmv_UA[i])) %>%
    filter(marker %in% sig.markers_by.cluster)
  
  #dataset_dummy_2$marker <- factor(dataset_dummy_2$marker, levels =c("CCR7","CD45RA","CD45RO",
                    #"GZMB","KI67","TBET",
                    #"CTLA4","LAG3","PD1","PDL1",
                    #"TIGIT", "TIM3"))
  
  if (length(sig.markers_by.cluster) > 0) {
  baseline_plot <- ggplot(dataset_dummy_1, aes(x= age65_status, y= mmi, col = age65_status)) + 
  scale_color_manual(values = color_pal) + 
  geom_boxplot(size = 0.75,
               fill = 'white') +
  geom_jitter(alpha = 0.75, pch=19, size=1.5, position=position_jitterdodge(dodge.width=1, jitter.width=.5)) +
  ggtitle(paste(cytof_cluster_grouped_list_rmv_UA[i])) +
  #xlab("Baseline") +
  ylab("MMI") +
  #theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16), axis.text.y.left = element_text(angle = 0), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position = 'none', axis.text.y = element_text(color="black", size=11, angle = 90, hjust = 0.5), axis.text.x = element_text(color="black",size=11, angle = 0), axis.line = element_line(colour = "black"), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14)) + 
  stat_compare_means(comparisons = list(c("no","yes")), aes(label = after_stat(p.signif)), bracket.size = 0.5, size = 4) +
  scale_x_discrete(labels = c("Young", "Aged")) +
  facet_wrap(~ `marker`, scales = "free", ncol = length(sig.markers_by.cluster)) +
  theme(axis.text.y = element_text(size = 13),
        axis.text.x = element_text(size = 13),
        axis.title.x = #element_text(size = 14), 
        element_blank(),
        axis.title.y = element_text(size = 14),
        plot.title = element_text(size = 15),
        strip.text = element_text(size = 14), 
        strip.background = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0.1)))
  
  print(baseline_plot)
  
  baseline_plot_list[[i]] <- baseline_plot
  
  on.treatment_plot <- ggplot(dataset_dummy_2, aes(x= age65_status, y= mmi, col = age65_status)) + 
  scale_color_manual(values = color_pal) + 
  geom_boxplot(size = 0.75,
               fill = 'white') +
  geom_jitter(alpha = 0.75, pch=19, size=1.5, position=position_jitterdodge(dodge.width=1, jitter.width=.5)) +
  ggtitle(paste(cytof_cluster_grouped_list_rmv_UA[i])) +
  #xlab("Baseline") +
  ylab("MMI") +
  #theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16), axis.text.y.left = element_text(angle = 0), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position = 'none', axis.text.y = element_text(color="black", size=11, angle = 90, hjust = 0.5), axis.text.x = element_text(color="black",size=11, angle = 0), axis.line = element_line(colour = "black"), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14)) + 
  stat_compare_means(comparisons = list(c("no","yes")), aes(label = after_stat(p.signif)), bracket.size = 0.5, size = 4) +
  scale_x_discrete(labels = c("Young", "Aged")) +
  facet_wrap(~ `marker`, scales = "free", ncol = length(sig.markers_by.cluster)) +
  theme(axis.text.y = element_text(size = 13),
        axis.text.x = element_text(size = 13),
        axis.title.x = #element_text(size = 14), 
        element_blank(),
        axis.title.y = element_text(size = 14),
        plot.title = element_text(size = 15),
        strip.text = element_text(size = 14), 
        strip.background = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0.1)))
  
  print(on.treatment_plot)
  
  on.treatment_plot_list[[i]] <- on.treatment_plot
  }
  
}
```

### Plot saving

```{r}
ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "B_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[1]], 
       height = 3, width = 6,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "B_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[1]], 
       height = 3, width = 6,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "DNT_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[2]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "DNT_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[2]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "NK_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[4]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "NK_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[4]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcCM_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[5]], 
       height = 3, width = 6,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcCM_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[5]], 
       height = 3, width = 6,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcEFF_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[6]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcEFF_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[6]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcEM_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[7]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcEM_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[7]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcN_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[8]], 
       height = 3, width = 18,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "TcN_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[8]], 
       height = 3, width = 18,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Th17_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[9]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Th17_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[9]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Th2_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[10]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Th2_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[10]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "ThCTL_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[13]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "ThCTL_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[13]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "ThN_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[14]], 
       height = 3, width = 4,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "ThN_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[14]], 
       height = 3, width = 4,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Treg_baseline_age65.status_cytof.pdf"), 
       plot = baseline_plot_list[[15]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)

ggsave(filename = paste0(output.path.pdf, "/", Sys.Date(), "_", "Treg_on.treatment_age65.status_cytof.pdf"), 
       plot = on.treatment_plot_list[[15]], 
       height = 3, width = 2,
       device = cairo_pdf,
       dpi = 600)
```

### Supporting data

### Supporting data file generation

```{r}
baseline_cytof.marker_supporting.data <- clinical_cytof.marker_baseline_v1 %>%
  select(`pub_id paper`,age65_status,immune_cell,marker,mmi) %>%
  mutate(age65_status = str_to_title(age65_status)) %>%
  rename(`Immune cluster` = "immune_cell",
         `Aged status` = "age65_status",
         `Mean metal intensity (MMI)` = "mmi",
         `Marker` = "marker",
         `Subject ID` = "pub_id paper")

on.treatment_cytof.marker_supporting.data <- clinical_cytof.marker_on.treatment_v1 %>%
  select(`pub_id paper`,age65_status,immune_cell,marker,mmi) %>%
  mutate(age65_status = str_to_title(age65_status)) %>%
  rename(`Immune cluster` = "immune_cell",
         `Aged status` = "age65_status",
         `Mean metal intensity (MMI)` = "mmi",
         `Marker` = "marker",
         `Subject ID` = "pub_id paper")

```

### Exporting 

```{r}

write.xlsx(baseline_cytof.marker_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "baseline_cytof.marker.analysis_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)

write.xlsx(on.treatment_cytof.marker_supporting.data, 
           file = here(output.path.excel, paste0(Sys.Date(), "_", "on.treatment_cytof.marker.analysis_supporting.data.xlsx")), colNames = TRUE, rowNames = FALSE, append = FALSE)

```

